### 2020-02-08 nlp_models.tex
updated main sections

\michael{WHERE: Say large $\alpha$ values don't mean much.}
We do note, however, that many NLP models, such as the GPT and BERT, may have weight matrices with unusually large PL exponents
($\alpha>6$), which indicates these matrices may be \emph{under}-correlated (i.e over-paramaterized), although the truncated PL fit
itself may not be very reliable because the MLE estimator is unreliable in this range.  Phenomenologically, if we examine
the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase\cite{MM}.
We have conjectured previously that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$, 
and improved versions of GPT (shown below, and BERT (not shown) confirm this.

\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
To avoid confusion, let us clarify the differences between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated power law (PL), parameterized by 2 values:
the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  (Technically, we also need the minimum
eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
The PL exponent $\alpha$ measures of the 
amount of correlation in a DNN layer weight matrix $\mathbf{W}$.  It is valid for $\lambda<\lambda_{max}$, and
it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ Norm.
$(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 
and lets us compute a balanced, weighted average log PL norm metrix for the entire DNN.  
The $\hat{\alpha}$ metric also behaves like an imrpoved, weighted log average Spectral Norm, and may track this metric in some cases.

>>>>

%\michael{WHERE: Say large $\alpha$ values don't mean much.}
\paragraph{What does large $\alpha$ mean?}
We do note, however, that many NLP models, such as the GPT and BERT, may have weight matrices with unusually large PL exponents
($\alpha\gg 6$), which indicates these matrices may be \emph{under}-correlated (i.e over-parameterized), although the truncated PL fit
itself may not be very reliable because the MLE estimator is unreliable in this range.  Phenomenologically, if we examine
the ESD visually, we can usually describe these $\mathbf{W}$ as in the \emph{Bulk-Decay} or \emph{Bulk-plus-Spikes} phase\cite{MM}.
We have conjectured previously that very well-trained DNNs would not have many \emph{outlier} $\alpha>6$, 
and improved versions of GPT (shown below, and BERT (not shown) confirm this.

%\michael{WHERE: Highlight difference between $\alpha$ and $\hat{\alpha}$.}
\paragraph{The differences between $\alpha$ and $\hat{\alpha}$.}
To avoid confusion, let us clarify the differences between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated power law (PL), parameterized by 2 values:
the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  (Technically, we also need the minimum
eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
The PL exponent $\alpha$ measures of the 
amount of correlation in a DNN layer weight matrix $\mathbf{W}$.  It is valid for $\lambda<\lambda_{max}$, and
it scale invariant (does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.)
The Weighted-Alpha metric, $\hat{\alpha}=\alpha\log\lambda_{max}$, approximates the log $\alpha$ Norm.
$(\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha})$ (derived using statistical mechanics and RMT \nred{(in progress)}, 
and lets us compute a balanced, weighted average log PL norm metrix for the entire DNN.  
The $\hat{\alpha}$ metric also behaves like an imrpoved, weighted log average Spectral Norm, and may track this metric in some cases.
